---
title: "Assignment 03 - Web Scraping and Text Mining"
author: "Amei Hao"
date: "10/14/2020"
output: html_document
---
The learning objectives are to conduct data scraping and perform text mining.
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(xml2)
library(httr)
library(readr)
library(tidytext)
library(ggplot2)
```

#APIs
Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

As we did in lab 7. Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).
```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")
# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/span")
# Turning it into text
counts <- as.character(counts)
# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```
There are 564 articles that related to sars-cov-2 trial vaccine.

```{r}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(db="pubmed",term="sars-cov-2 trial vaccine", retmax=250)
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)
# Turn the result into a character vector
ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]
# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")

```

```{r}
 publications <- GET(
   url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
   query = list(
     db = "pubmed",
     id = paste(ids,collapse = ","),
     retmex = 250,
     rettype = "abstract"
     )
 )
# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```


```{r}
# Form a dataset.
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
abstracts <- str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
abstracts <- str_replace_all(abstracts, "\\s+"," ")
#title them without an abstract
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+"," ")
#publication date
dates <- str_extract_all(pub_char_list, "<PubDate>(\\n|.)+</PubDate>")
dates <- str_remove_all(dates, "</?[[:alnum:]]+>")
dates <- str_replace_all(dates, "\\s+", " ")
#name of the journal
journal <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journal <- str_remove_all(journal, "</?[[:alnum:]]+>")
journal <- str_replace_all(journal, "\\s+", " ")
database <- data.frame(
   PubMedID = ids,
   Title = titles,
   Journal = journal,
   Date = dates,
   Abstract = abstracts
 )
 knitr::kable(database)
```


#Text Mining
A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.

Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?
Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.
Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”)
What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r}
# download data
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", "pubmed_data.csv", method="libcurl", timeout = 60)
newdata <- data.table::fread("pubmed_data.csv")
head(newdata)
```
##Tokenize the abstracts and count the number of each token
```{r}
library(tidytext)
library(tidyverse)
library(dplyr)
newdata %>%
   unnest_tokens(output = token,input = abstract) %>%
   count(token, sort = TRUE)
#tokenize & count
newdata %>%
   unnest_tokens(token,abstract)%>%
   anti_join(stop_words, by = c("token" = "word"))%>%
   count(token,sort = TRUE)

#remove the stop words
newdata %>%
  unnest_tokens(token,abstract)%>%
  anti_join(stop_words, by = c("token" = "word"))%>%
  count(token,sort = TRUE)

#get 5 most common tokens
newdata %>%
  unnest_tokens(token,abstract)%>%
  anti_join(stop_words, by = c("token" = "word"))%>%
  group_by(term)%>%
  count(token)%>%
  top_n(5,n)
```
There are 20,567 tokens. We have top five most common words: 'the'(n=28,126), 'of' (n=24,760), 'and'(n=19,993), 'in'(n=14,653), and 'to'(n=10,920).


After removing the stop words, we have 19,987 rows left. 
The most 5 common token is 'covid'(n=7,275),'19'(n=7,080), 'patients' (n=4,674), 'cancer'(n=3,999), and 'prostate'(n=3,832).

## Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.
```{r}
library(ggplot2)
library(tidytext)
library(tidyverse)
library(dplyr)

newdata %>%
  unnest_ngrams(token, abstract, n=2) %>%
  count(token, sort = TRUE) %>%
  top_n(10, n)%>%
  ggplot(aes(x = n, y = fct_reorder(token, n)))+
  geom_col()
```
# Calculate the TF-IDF value
```{r}
newdata %>%
  unnest_tokens(token, abstract) %>%
  group_by(term)%>%
  anti_join(stop_words, by = c("token" = "word"))%>%
  count(token, sort = TRUE)%>%
  bind_tf_idf(token, term, n) %>%
  top_n(5,n)%>%
  arrange(desc(tf_idf))
```
The 5 tokens from each search term with the highest TF-IDF value are as following: 
covid: covid, pandemic, coronavirus, sars, cov
prostate cancer: prostate, androgen, psa, prostatectomy, castration
preeclampsia: eclampsia, preeclampsia, pregnancy, maternal, gestational
meningitis: meningitis, meningeal, pachymeningitis, csf, meninges
cystic fibrosis: cf, fibrosis, cystic, cftr, sweat

covid:"covid" has the highest tf_idf value;
prostate cancer: "postate";
preeclampsia:"eclampsia";
meningitis:"meningitis";
cystic fibrosis:"cf"